Sea: A lightweight data-placement library for Big Data scientific computing

### 2022-03-30 paper/sea-comp/paper.tex
fixes to comp paper

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{longtable}
\usepackage[binary-units=true]{siunitx}
\usepackage{multirow}
\usepackage{subcaption}
%\usepackage{algorithm}
\usepackage[ruled,norelsize]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{tabularx}

@@

% \algnewcommand\algorithmicforeach{\textbf{for each}}
% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
% \algblock{Input}{EndInput} \algnotext{EndInput} \algblock{Output}{EndOutput}
% \algnotext{EndOutput} \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
#1}}}

@@

@@

\subsubsection{Benefits}
Sea removes the need to implement the logic to redirect data to the appropriate
storage locations. Furthermore, it implements logic to ensure that files are
written to the best possible location at any given time. It is both pipeline and
infrastructure agnostic.

@@

Glibc interception is achieved by writing wrappers to existing glibc functions.
Most importantly, every glibc function accepting a file path needs to be
wrapped. The wrappers take any input filepath that is located within the
user-provided Sea mountpoint and convert it a filepath pointing to the best
available storage device.

@@

Our tests have been executed on three different operating systems: Ubuntu
(xenial, bionic, focal), Fedora (32, 33, 34) and Centos (6, 7). Using this test
matrix, the Sea tests were able to span seven glibc versions ranging from 2.17
to 2.33. As glibc file system functions may change over time, Sea behaviour with
newer and older versions of glibc is undefined. 
% ldd (GNU libc) 2.17 ldd (Ubuntu GLIBC 2.23-0ubuntu11.3) 2.23 ldd (Ubuntu GLIBC
% 2.27-3ubuntu1.4) 2.27 ldd (GNU libc) 2.28 ldd (Ubuntu GLIBC 2.31-0ubuntu9.2)
% 2.31 ldd (GNU libc) 2.31 ldd (GNU libc) 2.32 ldd (GNU libc) 2.33

\subsubsection{Flushing and eviction modes}
\begin{figure*}

    \centering
    \includegraphics[scale=0.8]{figures/sea-flushevict.pdf}%
\caption{The four Sea flush and eviction modes }
\label{fig:sea-comp:fe}
\end{figure*}
%\todo{add diagrams here?}
Flushing and eviction mechanisms give way to four different modes: flush-only,
evict-only, flush-and-evict and keep. A flush-only is synonymous with a file
copy in that a file that already exists within a Sea cache will be copied to
persistent storage. This operation is particularly useful when a file will be
reused by the application, and thus benefits from remaining in cache, but also
needs to be communicated between other nodes or needs to be persisted for
post-processing analysis. To flush-only a file, the file must only be specified
in the \texttt{.sea\_flushlist}.

@@

@@

      Experimental results can be prone to error and variability, particularly
      in instances where there a many factors at play, such as in cluster
      environments. To improve confidence in our results, we developed a
      simplified performance model that could help describe the behaviours found
      in our experimental results. Since different parallel file systems may
      operate differently, our baseline model will be based on Lustre which is
      commonly used on HPC infrastructure.

      For data intensive use cases, the makespan models for both Lustre and Sea
      can be broken down into two components: The amount of time it takes read
      the data and the amount of time it takes to write the data. With more
      heterogeneous applications (some components are compute intensive whereas
      others are data intensive), a third component, consisting of compute time,
      can be added. Furthermore, latency may also play a significant role
      application makespan, particularly in scenarios with large amounts of
      small files. We choose to ignore latency costs in our model and make the
      assumption that the application bottleneck is the bandwidth, however, for
      more accurate estimates we might consider the addition of file system
      latency, as a fourth model component.

@@

      %\spacing{1.5}
      For the sake of simplicity, the above models assume that the network
      bandwidth between the compute and data nodes is the same. This, however,
      may not necessarily be the case. Furthermore, the model also assumes that
      each file can only be located on a single disk, meaning that the parallel
      bandwidth can at maximum be as fast as all Lustre disks combined and as
      slow as the minimum number of compute threads reading and writing files.

@@

      Page cache is difficult to summarize accurately and effectively within a
      model. For one, we must not only consider available memory and anonymous
      memory used by the application, but we must also consider which pages are
      candidates for eviction and which files they belong to. In addition, in
      the case of writes, we must consider asynchronous flushing and the
      throttling that may occur as a consequence of surpassing the
      \texttt{dirty\_ratio}. Furthermore, Lustre also has its own user-defined
      settings for how it interacts with the cache that would add additional
      complexities to the model. As a result, we assume two possible scenarios,
      one in which page cache is never used
      (Equation~\ref{eq:sea-comp:lustrenpc}) and one in which all application
      I/O occurs entirely within page cache, excluding the first read which must
      occur on Lustre (Equation~\ref{eq:sea-comp:lustrepc}). These two models
      allow us to define the bounds of Lustre's performance.

@@

      Here, $M_{n}$ represents the makespans of the different possible storage
      levels (e.g. tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
      makespan of an application reading and writing to Sea. For our model, we
      will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
      storage, and 3) slow parallel file system layer.

@@

      The local disk  makespan model is similar to the tmpfs makespan model,
      however, we must ensure to disregard any data that has already been
      written to tmpfs. Furthermore, in Sea, it is possible to leverage however
      many disk-based file systems are available for use ($g$). For our model,
      we assume that the size of each device is identical. The makespan model
      can be defined as follows:

@@

\subsubsection{Infrastructure}
      \begin{table}
      \centering
      \begin{tabular}{@{}|c|c|c|@{}}
       \hline
       Storage layer & Action & Avg. bandwidth (MiB/s) \\
              \hline
       \multirow{3}{*}{tmpfs} & read & 6676.48 \\
       & cached read & 6318.08  \\
       & write & 2560.00 \\
       \hline
       \multirow{3}{*}{local disk} & read & 501.70  \\
       & cached read & 7034.88 \\
       & write & 426.00 \\
       \hline
       \multirow{3}{*}{Lustre} & read & 1381.14 \\
       & cached read & 6103.04  \\
       & write & 121.00  \\

       \hline

      \end{tabular}
      \caption{Storage benchmarks}
      \label{table:sea-comp:fs}
      \end{table}
            Our experiments were executed on a Centos 8.1 (Linux kernel 4.18.0)
      cluster with 8 compute nodes, a 4 data node Lustre server with 1 dedicated
      metadata node. Each compute node is equipped it two Intel(R) Xeon(R) Gold
      6130 CPUs, \SI{250}{\gibi\byte} of memory with \SI{126}{\gibi\byte} of
      tmpfs space and 6 \SI{447}{\gibi\byte} Intel SSDSC2KG480G8R SSDs. The data
      nodes each contain 11 \SI{10}{\tera\byte} HGST HUH721010AL HDD OSTs and
      \SI{62}{\gibi\byte} memory. The metadata server contains a Toshiba
      KPM5XVUG960G \SI{960}{\giga\byte} MDT. The network bandwidth is
      \SI{25}{\giga\bit}E and uses TCP for communication. Jobs are scheduled on
      the cluster from a controller node using Slurm with cgroups. Swapping is
      disabled and Lustre dirty writes is limited to \SI{1}{\giga\byte} per OST.

@@

      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea with the
      in-memory configuration speeds up execution the majority of the
      conditions. The largest speedups observed can be seen in
      Figure~\ref{fig:sea-comp:processes} at 32 processes, where the speedup is
      nearly 3$\times$. We believe that the speedup here is primarily due to the
      contention on Lustre. While each local disk has an average of 5 processes
      trying to write concurrently to it at 32 processes, each Lustre has around
      3 concurrent processes while to it. Whereas one might assume this means
      Lustre would have less contention overall and should perform better, there
      are significant bandwidth differences to account for between the local
      disks and Lustre (see Table~\ref{table:sea-comp:fs}) that would result in
      improved local disk performance despite increased contention. Furthermore,
      Lustre has a centralized metadata server that is responsible for
      determining which OST is assigned to each block, guaranteeing a certain
      amount of load-balance on the storage disks. This is in contrast with the
      local disks, who do not rely on a metadata server and are selected by Sea
      via a random shuffling.

@@

      When varying the number of nodes (Figure~\ref{fig:sea-comp:nodes}), we
      achieved the greatest speedup at 5 nodes, with a speedup of 2.4$\times$.
      Similar to our experiments with multiprocessing, the speedup here appears
      to be due to increased contention on Lustre. However, there is one main
      difference here, and that is that only Sea is experiencing increased
      contention, as the contention within the compute nodes is fixed. Both the
      model and experimental results state that the speedup is approaching a
      plateau, however, it is expected that once the number of threads writing
      to Lustre exceeds the number of OSTs (e.g. 9 nodes), we will observe an
      even greater speedup from Sea due to the increased contention on Lustre.

@@

      \subsection{Model accurately predicts performance trends}

      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model correctly
      describes The bounds for all but two experiments: Experiment 3
      (Figure~\ref{fig:sea-comp:iterations}) and Experiment 4
      (Figure~\ref{fig:sea-comp:processes}). In Experiment 3, the model
      incorrectly represents the bounds for 1 iteration. Since, in this case,
      all the data should be able to fit in page cache for both Sea and Lustre,
      it is possible that memory bandwidth has been overestimated, resulting in
      incorrect model predictions.

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{longtable}
\usepackage[binary-units=true]{siunitx}
\usepackage{multirow}
\usepackage{subcaption}
%\usepackage{algorithm}
\usepackage[ruled,norelsize]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{tabularx}

@@

\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
% \algnewcommand\algorithmicforeach{\textbf{for each}}
% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
% \algblock{Input}{EndInput} \algnotext{EndInput} \algblock{Output}{EndOutput}
% \algnotext{EndOutput} \newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\newcommand{\todo}[1]{\marginpar{\parbox{18mm}{\flushleft\tiny\color{red}\textbf{TODO}:
#1}}}

@@

Sea removes the need to implement the logic to redirect data to the appropriate
storage locations. Furthermore, it implements logic to ensure that files are
written to the best possible location at any given time. It is both pipeline and
infrastructure agnostic.

@@

@@

Glibc interception is achieved by writing wrappers to existing glibc functions.
Most importantly, every glibc function containing accepting a file path as input parameter needs to be
wrapped. The wrappers take any input filepath that is located within the
user-provided Sea mountpoint and convert it a filepath pointing to the best
available storage device.

@@

Our tests have been executed on three different operating systems: Ubuntu
(xenial, bionic, focal), Fedora (32, 33, 34) and Centos (6, 7). Using this test
matrix, the Sea tests were able to span seven glibc versions ranging from 2.17
to 2.33. As glibc file system functions may change over time, Sea behaviour with
newer and older versions of glibc is unknown. 
% ldd (GNU libc) 2.17 ldd (Ubuntu GLIBC 2.23-0ubuntu11.3) 2.23 ldd (Ubuntu GLIBC
% 2.27-3ubuntu1.4) 2.27 ldd (GNU libc) 2.28 ldd (Ubuntu GLIBC 2.31-0ubuntu9.2)
% 2.31 ldd (GNU libc) 2.31 ldd (GNU libc) 2.32 ldd (GNU libc) 2.33

\subsubsection{Memory-management}
      \begin{table}
      \centering
      \begin{tabular}{ccc}
        \toprule
       Mode & .sea\_flushlist & .sea\_evictlist \\
       \midrule
       Copy & yes & no \\
       Remove & no & yes \\
       Move & yes & yes \\
       Keep & no & no \\

       \bottomrule

      \end{tabular}
      \caption{Sea's memory management modes}
      \label{table:sea-comp:fe}
      \end{table}
%\todo{add diagrams here?}
Memory-management is an integral component of Sea as deallocating data from cache
makes room for future files to be stored within the higher-performance storage devices.
Generalizing an efficient memory-management strategy across all applications is a challenging task.
Since Sea works on a per-application basis, memory-management in Sea is application-specific, configured
via two files: \texttt{.sea\_flushlist} and \texttt{.sea\_evictlist}.

Memory management in Sea consists of four different modes, as described in Table~\ref{table:sea-comp:fe}: flush-only,
evict-only, flush-and-evict and keep. A flush-only is synonymous with a file
copy in that a file that already exists within a Sea cache will be copied to
persistent storage. This operation is particularly useful when a file will be
reused by the application, and thus benefits from remaining in cache, but also
needs to be communicated between other nodes or needs to be persisted for
post-processing analysis. To flush-only a file, the file must only be specified
in the \texttt{.sea\_flushlist}.

@@

A third text file, namely the \texttt{.sea\_prefetchlist}, exists that can be populated by input
files to be prefetched. At this time, Sea cannot determine when prefetched files are no longer needed by the pipeline, and
therefore should not be evicted. For files to be prefetched, they must be located within Sea's mountpoint at startup.

@@

      Experimental results can be prone to error and variability, particularly
      in instances where there a many factors at play, such as in distributed cluster
      environments. To improve confidence in our results, we developed a
      simplified performance model that could help describe the behaviours observed in our experimental
      results. Since different PFS may
      operate differently, our baseline model will be based on Lustre which is
      commonly used on HPC infrastructure.

      For data-intensive use cases, the makespan models for both Lustre and Sea
      can be broken down into two components: The amount of time it takes read
      the data and the amount of time it takes to write the data. With more
      heterogeneous applications (some components are compute-intensive whereas
      others are data-intensive), a third component, consisting of compute time,
      can be added. Furthermore, latency may also play a significant role
      application makespan, particularly in scenarios with large amounts of
      small files. We choose to ignore latency costs in our model and make the
      assumption that the application bottleneck is the bandwidth, however, for
      more accurate estimates we might consider the addition of file system
      latency, as a fourth model component.

@@

      %\spacing{1.5}
      For the sake of simplicity, the above models assume that the network
      bandwidth between the compute and data nodes is the same. However, this
      may not necessarily be the case. Furthermore, the model also assumes that
      each file can only be located on a single disk, meaning that the parallel
      bandwidth can at maximum be as fast as all Lustre disks combined and as
      slow as the minimum number of compute threads reading and writing files.

@@

      Page cache is difficult to summarize accurately within a
      model. For one, we must not only consider available memory and anonymous
      memory used by the application, but we must also consider which pages are
      candidates for eviction and which files they belong to. In addition, in
      the case of writes, we must consider asynchronous flushing and the
      throttling that may occur as a consequence of surpassing the
      dirty\_ratio. Furthermore, Lustre also has its own user-defined
      settings for how it interacts with the cache that would add additional
      complexities to the model. As a result, we assume two possible scenarios,
      one in which page cache is never used
      (Equation~\ref{eq:sea-comp:lustrenpc}) and one in which all application
      I/O occurs entirely within page cache, excluding the first read which must
      occur on Lustre (Equation~\ref{eq:sea-comp:lustrepc}). These two models
      allow us to define the bounds of Lustre's performance.

@@

      Here, $M_{n}$ represents the makespans of the different possible storage
      levels (e.g. tmpfs, NVMe, SSD, HDD, Lustre) and $M_{S}$ represent the
      makespan of an application reading and writing to Sea. For our model, we
      will assume 3 storage layers: 1) fast tmpfs, 2) intermediate local SSD
      storage, and 3) slow parallel file system layer. It is important to note that
      the model assumes that I/O never occurs in distinct storage layers in parallel.
      Although Sea tries to exhaust the layers, starting from the fastest, it is Nevertheless
      possible that Sea writes to two storage layers at the same time, for instance, when one becomes full
      and another parallel thread needs to perform a write.

@@

      The local disk  makespan model is similar to the tmpfs makespan model,
      although we must disregard any data that has already been
      written to tmpfs. Furthermore, in Sea, it is possible to leverage however
      many disk-based file systems are available for use ($g$). For our model,
      we assume that the size of each device is identical. The makespan model
      can be defined as follows:

@@

\subsubsection{Infrastructure}
      \begin{table}
      \centering
      \begin{tabular}{ccc}
       \toprule
       Storage layer & Action & Avg. bandwidth (MiB/s) \\
       \midrule
      \multirow{3}{*}{tmpfs} & read & 6676.48 \\
      & cached read & 6318.08  \\
      & write & 2560.00 \\
      \midrule
       \multirow{3}{*}{local disk} & read & 501.70  \\
       & cached read & 7034.88 \\
       & write & 426.00 \\
       \midrule
       \multirow{3}{*}{Lustre} & read & 1381.14 \\
       & cached read & 6103.04  \\
       & write & 121.00  \\

       \bottomrule

      \end{tabular}
      \caption{Storage benchmarks}
      \label{table:sea-comp:fs}
      \end{table}
            Our experiments were executed on a Centos 8.1 (Linux kernel 4.18.0)
      cluster with 8 compute nodes, a 4 data node Lustre server with 1
      metadata node. Each compute node is equipped it two Intel(R) Xeon(R) Gold
      6130 CPUs, \SI{250}{\gibi\byte} of memory with \SI{126}{\gibi\byte} of
      tmpfs space and 6 \SI{447}{\gibi\byte} Intel SSDSC2KG480G8R SSDs. The data
      nodes each contain 11 \SI{10}{\tera\byte} HGST HUH721010AL HDD OSTs and
      \SI{62}{\gibi\byte} memory. The metadata server contains a Toshiba
      KPM5XVUG960G \SI{960}{\giga\byte} MDT. The network bandwidth is
      \SI{25}{\giga\bit}E and uses TCP for communication. Jobs are scheduled on
      the cluster from a controller node using Slurm with cgroups. Swapping is
      disabled and Lustre dirty writes is limited to \SI{1}{\giga\byte} per OST.

      There was no background load on the cluster for the duration of the experiments.

@@

      As can be seen in Figure~\ref{fig:sea-comp:benchmarks}, Sea using
      in-memory configuration speeds up execution the majority of the
      conditions. The largest speedups observed can be seen in
      Figure~\ref{fig:sea-comp:processes} at 32 processes, where the speedup is
      nearly 3$\times$. We believe that the speedup is primarily due to the
      contention on Lustre. While each local disk has an average of 5 processes
      trying to write concurrently to it at 32 processes, each Lustre has around
      3 concurrent processes while to it. Whereas one might assume this means
      Lustre would have less contention overall and should perform better, there
      are significant bandwidth differences to account for between the local
      disks and Lustre (see Table~\ref{table:sea-comp:fs}) that would result in
      improved local disk performance despite increased contention. Furthermore,
      Lustre has a centralized metadata server that is responsible for
      determining which OST is assigned to each block, guaranteeing a certain
      amount of load-balance on the storage disks. This is in contrast with the
      local disks, who do not rely on a metadata server and are selected by Sea
      via a random shuffling.

@@

      When varying the number of nodes (Figure~\ref{fig:sea-comp:nodes}), we
      achieved the greatest speedup at 5 nodes, with a speedup of 2.4$\times$.
      Similar to our experiments with multiprocessing, the speedup appears
      to be due to increased contention on Lustre. However, there is one main
      difference, and that is that only Sea is experiencing increased
      contention, as the contention within the compute nodes is fixed. Both the
      model and experimental results state that the speedup is approaching a
      plateau, however, it is expected that once the number of threads writing
      to Lustre exceeds the number of OSTs (e.g. 9 nodes), we will observe an
      even greater speedup from Sea due to the increased contention on Lustre.

@@

      \subsection{Performance trends described by the model}

      Figure~\ref{fig:sea-comp:benchmarks} shows us that our model can
      describe the performance trends for all but two experiments: Experiment 3
      (Figure~\ref{fig:sea-comp:iterations}) and Experiment 4
      (Figure~\ref{fig:sea-comp:processes}). In Experiment 3, the model
      incorrectly represents the bounds for 1 iteration. Since, in this case,
      all the data should be able to fit in page cache for both Sea and Lustre,
      it is possible that memory bandwidth has been overestimated, resulting in
      incorrect model predictions.
